|

### 2022-10-04 paper_comparison/paper_comparison/paper_comparison.Rmd
intro done

```{r poly-conventional-simulation}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

@@

```{r heter-conventional-simulation}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

@@

Diagnosing a model is the key to determining whether there is anything importantly wrong. For linear regression analysis, it is typical to interrogate the residuals. Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong. There are many ways that residuals could be assessed.

Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved. But if the distribution is skewed, bimodal, multimodal, or contains outliers, there is cause for concern. The distribution could also be inspected by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].

Plotting the residuals against predicted values and each of the explanatory variables on a scatter plot is a recommend way to scrutinize their relationships. If there is any visually discoverable patterns, the model is potentially misspecified. However, it is a very difficult task for a human judge, though to make a decision that there's nothing there. It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none. Generally, one looks for departures from nothingness like non-linear dependency or heteroskedasticity. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

There is an abundance of literature describing appropriate diagnostic methods for linear regression: @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals. The diligent reader of these sage writings will also notice sentences that express sentiments like *based on their experience, statistical tests are not widely used in regression diagnostics. The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.* There is a common guidance by experts that plots are the best for diagnosing model fits.

This is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next background section describes the the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers. Section \ref{experimental-design} describes the experimental setup to enable a comparison between decision made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section \ref{results}. We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.

@@

```{r residual-plot-common-departures, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0. The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}

@@

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \ref{fig:residual-plot-common-departures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot.

Non-linearity, heterskedasticity and non-normality are perhaps the three mostly commonly checked departures. Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. An example residual plot containing visual pattern of non-linearity is given at the top-right of Figure [ref here]. One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model. Heterskedasticity refers to the presence of nonconst error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values. Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:residual-plot-common-departures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis. Compared to non-linearity and heterskedasticity, non-normality is usually harder to be detected from a residual plot since scatter plot is not excel in revealing marginal distribution. A favourable graphical summarise for this task is the quantile-quantile plot. However, for a consistent comparison in the later part of this paper, residual plot will be the focus. Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model. In the case that the normality assumption is violated, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line. For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:residual-plot-common-departures}.

@@

Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding model structural test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}. In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures that they do not intend to examine. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999]. This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.

@@

  table_dat %>%
  knitr::kable(col.names = c("Plot", "Departures", "Model structural", "Breusch-Pagan", "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" )

@@

<!-- TODO:Clarify this is only for integer c_i -->

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},
\end{equation}
```

@@

<!-- TODO:Introduce the concept of c interesting plot -->

@@

As discussed in @majumder_validation_2013, individual's skill will affect the number of observers who identify the actual data plot from the lineup. Thus, the power of a visual test depends on the subject-specific abilities. Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013]. However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation. Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor. The same model was applied to different subsets of the survey data for making comparison of the power.

@@

## Experiment overview

@@

Three experiments were conducted to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics. The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test. Meanwhile, the experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test. The experiment III is designed for collecting human responses to lineup with only good residual plots such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated. Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups performed by 443 subjects throughout three experiment.

@@

Two types of departures, namely non-linearity and heteroskedasticity, were considered with the corresponding data generating process being designed for experiment I and II.

@@

```{r results='asis'}
data.frame(j = c(2, 3, 6, 18),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
  knitr::kable(col.names = NULL,
               align = "rl",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Parameter values for $j$ and $X_{raw}$', 
               label = "model-parameter-table" ) %>%
  kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", "Distribution of sym4\n")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym4", "$X_{raw}$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

The values of $j$ was chosen so that distinct shapes of non-linearity were included in the residual plot. A greater value of $j$ will result in a curve with more turning points. As shown in Figure \ref{fig:different-shape-of-herimite}, it includes "U" shape, "S" shape, "M" shape and "Triple-U" shape. It is expected that the "U" shape will be the easiest one to detect because complex shape tends to be concealed by cluster of data points.

@@

Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

@@

Figure \ref{fig:example-poly-lineup} shows one of the lineups used in experiment I. This lineup was produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location was four. All five subjects correctly identified the actual data plot for this lineup.

```{r example-poly-lineup, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

@@

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("a = -1: Left-triangle shape")) +
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("a = 0: Butterfly shape")) +
  theme_light() -> p2

VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("a = 1: Right-triangle shape")) +
  theme_light() -> p3

patchwork::wrap_plots(p1, p2, p3, ncol = 2)
```

```{r example-heter-lineup, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
VI_MODEL$plot_lineup(vi_lineup[["heter_169"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

An example lineup of this model is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$. The actual data plot location was 15. 8 out of 11 subjects correctly identified the actual data plot for this lineup.

@@

### Controlling the strength of the signal

As summarised by Table \ref{tab:model-parameter-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.

Three different sample sizes were used (n = 50, 100, 300) in all three experiments. It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.

```{r different-n, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(n~"="~300)) + 
  theme_light() -> p1

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
  ggtitle(quote(n~"="~50)) + 
  theme_light() -> p2

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
  ggtitle(quote(n~"="~100)) + 
  theme_light() -> p3

patchwork::wrap_plots(p2, p3, p1, ncol = 2)
```

```{r different-sigma, fig.cap="Example residual plots of the linear regression model used in experiment I. Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

```{r different-b, fig.cap="Example residual plots of the linear regression model used in experiment II. Five different values of $b$ are included in the experiment to control the strength of the signal."}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

ori_dat <- heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 0.25)$gen(300)

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("b = 0.25")) +
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 1)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 1")) +
  theme_light() -> p2

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 4)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 4")) +
  theme_light() -> p3

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 16)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 16")) +
  theme_light() -> p4

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 64)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 64")) +
  theme_light() -> p5

patchwork::wrap_plots(p1, p2, p3, p4, p5, ncol = 3)
```

@@

Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 443 different lineups. For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot.

In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.

Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.

>>>>

```{r}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

@@

```{r}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

@@

Diagnosing a model is the key to determining whether there is anything importantly wrong. For linear regression analysis, we typically interrogate the residuals for model diagnostics. Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.

We can assess residuals in multiple ways. Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved. But if the distribution is skewed, bimodal, multimodal, or contains outliers, there is cause for concern. The distribution could also be inspected by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].

Plotting the residuals against predicted values and each of the explanatory variables on a scatter plot is a recommend way to scrutinize their relationships. If there is any visually discoverable patterns, the model is potentially misspecified. However, correctly judging a residual plot where no pattern exists is a painstakingly difficult task for humans. It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none. [ET: I don't know if the former statement is true?] Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

Abundance of literature describe appropriate diagnostic methods for linear regression: @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals. The diligent reader of these sage writings will also notice sentences that express sentiments like *based on their experience, statistical tests are not widely used in regression diagnostics. The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.* A common guidance by experts is that optimal method for diagnosing model fits is by plotting the data.

The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next background section describes the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers [ET: CITATION HERE?]. Section \ref{experimental-design} describes the experimental setup to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section \ref{results}. We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.

@@

```{r commondepartures, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0. The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}

@@

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot. Two most commonly checked departures - non-linearity, heterskedasticity and non-normality will be discussed in this paper.

Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. An example residual plot containing visual pattern of non-linearity is given at the top-right of Figure [ref here]. One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.

Heterskedasticity refers to the presence of nonconst error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values. Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.

Compared to non-linearity and heterskedasticity, non-normality is usually harder to be detected from a residual plot since scatter plot is not excel in revealing marginal distribution. A favourable graphical summarise for this task is the quantile-quantile plot. However, for a consistent comparison in the later part of this paper, residual plot will be the focus. Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model. In the case that the normality assumption is violated, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line. For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:commondepartures}.

@@

Example residual plots given in Figure \ref{fig:commondepartures} are examined by the corresponding model structural test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}. In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures that they do not intend to examine. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999]. This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.

@@

  table_dat %>%
  knitr::kable(col.names = c("Plot", "Departures", "Model structural", "Breusch-Pagan", "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:commondepartures}. Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" )

@@

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \int_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
\end{equation}
```

@@

@@

As discussed in @majumder_validation_2013, individual's skill will affect the number of observers who identify the actual data plot from the lineup. Thus, the power of a visual test depends on the subject-specific abilities. Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013]. However, knowing this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject, and hence, the p-value can't be computed.

(1) Instead, we modelled the probability of a lineup being rejected directly using a logistic regression with the natural logarithm of the effect size as the only regressor. The same model was applied to different subsets of the survey data for making comparison of the power.

(2) Instead, we modelled the plot selection $c_i$ for the actual data plot $i$.

<!-- -   Very briefly explain what a lineup test is -->

<!-- -   Lineup (maybe using your previous residual plot for nonlinearity) -->

<!-- -   $p$-value calulcation, briefly, including how multiple selections are handled -->

<!-- -   power calculation, briefly, including how multiple selections are handled -->

```{=html}
<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009]. Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null. For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives. \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
-->
```

@@

## Overview

@@

Three experiments were conducted to investigate the difference between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics. The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test. The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test. The experiment III is designed for collecting human responses to lineup with only good residual plots such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated. Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups performed by 443 subjects throughout three experiment.

@@

Two types of departures were considered with the corresponding data generating process being designed for experiment I and II.

@@

```{r results='asis', eval = FALSE}
data.frame(j = c(2, 3, 6, 18),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
  knitr::kable(col.names = NULL,
               align = "rl",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Parameter values for $j$ and $X_{raw}$', 
               label = "parameter-table" ) %>%
  kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", "Distribution of sym4\n")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym4", "$X_{raw}$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot. As shown in Figure \ref{fig:different-shape-of-herimite}, these include "U" shape, "S" shape, "M" shape and "Triple-U" shape.

@@

<!-- The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. The example residual plots are given in Figure \ref{fig:different-sigma}. -->

```{r different-sigma, eval = FALSE, fig.cap="Example residual plots of the linear regression model used in experiment I. Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density. And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

@@

Figure \ref{fig:example-lineup} shows one of the lineups used in experiment I. This lineup was produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location was four. All five subjects correctly identified the actual data plot for this lineup.

```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

@@

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("Left-triangle shape")) +
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("Butterfly shape")) +
  theme_light() -> p2

VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("Right-triangle shape")) +
  theme_light() -> p3

patchwork::wrap_plots(p1, p2, p3, ncol = 3)
```

@@

### Treatments

To control the strength of the signal, two parameters were used including $\sigma$ and $b$ for controlling the standard deviation of the error term in experiment I and II respectively.

<!-- > Experiment I -->

<!-- -   Treatments for nonlinearity experiment (plots summarising levels) -->

<!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$. For this lineup, x out of k observers picked the actual data plot. Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value. -->

<!-- > Experiment II -->

<!-- -   Treatments for heteroskedasticity experiment (plots summarising levels) -->

@@
